Nested Sampling with Multiple Objective Functions

### 2018-06-03 doc/ms.tex
References

Nested Sampling \citep[NS][]{skilling} is popular general
Monte Carlo algorithm that can solve a wide range of problems in Bayesian
computation and statistical mechanics
\citep{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
Its key strength its ability to cope with phase transitions which cause
problems for many other methods \citep{skilling}.
In a Bayesian inference problem with unknown parameters denoted collectively
by a vector $\xx$, the
posterior distribution for the parameters given a data proposition $\data$ and
a `prior information' proposition $\II$ is
information $\II$) is:
\begin{eqnarray}
p(\xx | \data, \II) &=&
\frac{p(\xx | \II)p(\data | \xx, \II)}{p(\data | \II)}\\
&=& \frac{\pi(\xx)L(\xx)}{Z}
\end{eqnarray}
where $\pi(\xx)$ is the prior distribution, $L(\xx)$ is the likelihood
function, and $Z$ is the normalising constant, usually known as the
`marginal likelihood' or the `evidence':
\begin{eqnarray}
Z &=& \int \pi(\xx) L(\xx) \, \dx.\label{eqn:evidence}
\end{eqnarray}

$Z$ is useful because it is the (prior) probability of the data given that
$\xx$ is {\em one of}
the possibilities in the space being integrated over, and thus plays the role
of a likelihood for the parameter space as a whole.
The main goal of NS is to approximate $Z$, but it can also be used to make
Monte Carlo approximations of the posterior distribution, and to
calculate the `information',
or Kullback-Leibler divergence from the prior to the posterior:
\begin{eqnarray}
H &=& \int p(\xx | \data, \II) \ln
\left[\frac{p(\xx | \data, \II)}{p(\xx | \II)}\right] \, d\xx \\
&=& \int \frac{\pi(\xx) L(\xx)}{Z} \ln
\left[\frac{L(\xx)}{Z}\right] \, d\xx.
\end{eqnarray}
$H$ quantifies how compressed the posterior distribution is with
respect to the prior. For example, $H = 10$ implies, loosely speaking,
that the posterior occupies a fraction $e^{-10}$ of the prior mass.
$H$ can be interpreted quite literally as a measure of how much
the specific data $\data$ resolved the question ``what is the value
of $\xx$, precisely?''.

@@

\begin{thebibliography}{999} % if there are less than 10 entries, enter a one digit number

\bibitem[Baldock et al.(2015)]{2015arXiv150303404B} Baldock, R.~J.~N., 
P{\'a}rtay, L.~v.~B., Bart{\'o}k, A.~P., Payne, M.~C., Cs{\'a}nyi, G.\ 
2015.\ Determining pressure-temperature phase diagrams of materials.\ ArXiv 
e-prints arXiv: 1503.03404.

\bibitem[Brewer and Donovan(2015)]{exoplanet} Brewer, B.~J., 
Donovan, C.~P.\ 2015.\ Fast Bayesian inference for exoplanet discovery in 
radial velocity data.\ Monthly Notices of the Royal Astronomical Society 
448, 3206-3214.

\bibitem[Brewer(2014)]{rjobject} Brewer, B.~J.\ 2014.\ Inference 
for Trans-dimensional Bayesian Models with Diffusive Nested Sampling.\ 
ArXiv e-prints arXiv:1411.3921.

\bibitem[\protect\citeauthoryear{Brewer, P{\'a}rtay,
\& Cs{\'a}nyi}{2011}]{dnest} Brewer B.~J., P{\'a}rtay L.~B., Cs{\'a}nyi G., 2011,
Statistics and Computing, 21, 4, 649-656. arXiv:0912.2380

\bibitem[Cameron and Pettitt(2014)]{cameron} Cameron, E., and Pettitt, A., 2014,
``Recursive pathways to marginal likelihood estimation with
prior-sensitivity analysis.'' Statistical Science 29.3 (2014): 397-419.

\bibitem[Caticha(2008)]{caticha} Caticha, A.\ 2008.\ Lectures 
on Probability, Entropy, and Statistical Physics.\ ArXiv e-prints 
arXiv:0808.0012.

\bibitem[\protect\citeauthoryear{Gregory}{2005}]{gregory}
Gregory, Phil. Bayesian Logical Data Analysis for the Physical Sciences: A Comparative Approach with Mathematica® Support. Cambridge University Press, 2005.

\bibitem[Gull and Daniell(1978)]{gull} Gull, Stephen F., and Daniell, G., 1978
Image reconstruction from incomplete and noisy data. Nature (1978): 686-690.

\bibitem[Hansmann(1997)]{pt} Hansmann, Ulrich HE., 1997, ``Parallel tempering algorithm for conformational studies of biological molecules.'', Chemical Physics Letters 281, no. 1 (1997): 140-150.

\bibitem[Knuth and Skilling(2010)]{knuth_skilling} Knuth, K.~H., 
Skilling, J.\ 2010.\ Foundations of Inference.\
Axioms 1.1 (2012): 38-73.\
ArXiv e-prints arXiv:1008.4831.

\bibitem[Murray(2012)]{murray} Murray, Iain, Ghahramani, Zoubin, and MacKay, David.
``MCMC for doubly-intractable distributions.'' arXiv preprint arXiv:1206.6848 (2012).

\bibitem[P{\'a}rtay et al.(2014)]{2014PhRvE..89b2302P} P{\'a}rtay, L.~B., 
Bart{\'o}k, A.~P., Cs{\'a}nyi, G.\ 2014.\ Nested sampling for materials: 
The case of hard spheres.\ Physical Review E 89, 022302.

\bibitem[P{\'a}rtay et al.(2009)]{2009arXiv0906.3544P} P{\'a}rtay, L.~B., 
Bart{\'o}k, A.~P., Cs{\'a}nyi, G.\ 2009.\ Efficient sampling of atomic 
configurational spaces.\ ArXiv e-prints arXiv: 0906.3544.

\bibitem[Polson and Scott(2014)]{vertical}
Polson, Nicholas G., and James G. Scott. ``Vertical-likelihood Monte Carlo
Integration.'' arXiv preprint arXiv:1409.3601 (2014).

\bibitem[\protect\citeauthoryear{Skilling}{1998}]{massinf}
Skilling J., 1998, Massive Inference and Maximum Entropy, in Maximum Entropy
and Bayesian Methods, Kluwer Academic Publishers, Dordrecht/Boston/London p.14

\bibitem[\protect\citeauthoryear{Skilling}{2006}]{skilling} Skilling, J., 2006, Nested Sampling for General Bayesian Computation, Bayesian Analysis 4, pp. 833-860.

\bibitem[Vardi (1985)]{vardi}
Vardi, Yehuda. ``Empirical distributions in selection bias models.'', The Annals of Statistics (1985): 178-203.

\bibitem[Walter(2015)]{walter}
Walter, C.\ Point Process-based Monte Carlo estimation.\ arXiv: 1412.6368.
%% Reference 1
%\bibitem{ref-journal}
%Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.

%% Reference 2
%\bibitem{ref-book}
%Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.

\end{thebibliography}

>>>>

Nested Sampling \citep[NS][]{skilling} is general
Monte Carlo algorithm that can solve a wide range of problems in Bayesian
computation and statistical mechanics
\citep{2009arXiv0906.3544P, 2014PhRvE..89b2302P, 2015arXiv150303404B}.
Its key strength its ability to cope with phase transitions and other
features which cause
problems for many other methods.
In a Bayesian inference problem with unknown parameters denoted collectively
by a vector $\xx$, the
posterior distribution for the parameters given a data proposition $\data$ and
a `prior information' proposition $\II$ is:
\begin{eqnarray}
p(\xx | \data, \II) &=&
\frac{p(\xx | \II)p(\data | \xx, \II)}{p(\data | \II)}\\
&=& \frac{\pi(\xx)L(\xx)}{Z}
\end{eqnarray}
where $\pi(\xx)$ is the prior distribution, $L(\xx)$ is the likelihood
function, and $Z$ is the normalising constant, known as the
`marginal likelihood' or `evidence':
\begin{eqnarray}
Z &=& \int \pi(\xx) L(\xx) \, \dx.\label{eqn:evidence}
\end{eqnarray}

$Z$ is useful because it is the prior probability of the data conditional on
the hypothesis that $\xx$ is {\em one of}
the possibilities in the space being integrated over, along with
whatever additional assumptions were made. $Z$ thus plays the role
of a likelihood for the parameter space as a whole, and is a key input in
calculating the posterior probabilities of competing `models' (mutually
exclusive choices of $\II$).

The main goal of NS is to compute $Z$, but it can also be used to make
Monte Carlo approximations of the posterior distribution, and to
calculate the `information',
or Kullback-Leibler divergence from the prior to the posterior:
\begin{eqnarray}
H &=& \int p(\xx | \data, \II) \log
\left(\frac{p(\xx | \data, \II)}{p(\xx | \II)}\right) \, d\xx \\
&=& \int \frac{\pi(\xx) L(\xx)}{Z} \log
\left(\frac{L(\xx)}{Z}\right) \, d\xx.
\end{eqnarray}
$H$ quantifies how compressed the posterior distribution is with
respect to the prior. For example, $H = 100$ nats (the units obtained when
the logarithm is the natural logarithm) implies, loosely speaking,
that the posterior occupies a fraction $e^{-100}$ of the prior mass.
$H$ can be interpreted quite literally as a measure of how much
the specific data $\data$ resolved the question ``what is the value
of $\xx$, precisely?'' \citep{knuth, van_erp}.

@@

\bibliographystyle{mdpi}
\bibliography{references/references}
